Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-03-27 document/ms.tex
notes to self

STUFF HERE ABOUT EARTH ANALOGS.

This study is novel in a number of ways.
We create probabilistic information about the full exoplanet population
\emph{not} by weighting the detected objects by the inverse detection
efficiency (\citealt{dong, petigura}), but rather by forward modeling the
observed distribution subject to the detection efficiency.
Re-weighting the data by inverse selection probability---a method called
V-max in the quasar and galaxy luminosity function contexts---is not wrong,
but it produces a higher-variance estimate of the population than a justified
likelihood approach.
One simple intuition along these lines is that the inverse probability weights
can be large in dynamic range, so the effective number of samples, after
weighting, can be much smaller than the actual number of samples.
Another is that it tends to upweight most strongly the most marginal
detections in the data sample, since these tend to be the samples that lie in
the region of lowest completeness.  In this work, we build a justified
likelihood of the observed data, taking into account the completeness function
or detection efficiency.
Our likelihood function is a variable-rate Poisson likelihood; it makes the
fundamental assumption that the data points are independently drawn from the
model, but is agnostic about all other properties of the data or distribution
functions.
Indeed, the assumptions behind this study are weaker than those of any
previous study of the radius and period distribution of the Earth-analog
population (\citealt{dong, petigura}).

@@

We have made the first measurement of the Earth-like exoplanet population that
is based on forward modeling, takes proper account of the observational
uncertainties on planet radii, and makes use of extremely flexible
non-separable distributions for radius and period.
We find BLAH BLAH BLAH.

>>>>

\hoggtodo{
\kepler\ has been busting out exoplanets (CITE).
Many of them even appear possibly rocky and habitable (CITE).
Nonetheless, none yet known are true Earth analogs.
}

\hoggtodo{
This hasn't stopped anyone from estimating the rate at which Sun-like stars
host Earth-like planets on year-ish period orbits (CITE).
All of these rate estimates are extrapolations of one kind or another.
Here we perform an extremely conservative extrapolation of this kind.
}

\hoggtodo{
What would make an extrapolation like this conservative?
Using \emph{huge bins} does not a conservative estimate make.
You don't want to assume separabilities you can't justify.
You want to explore an enormous function space for the exoplanet rates or pdf.
We are going to do be conservative in these senses here, far more conservative
than anyone who has come before.
}

This study is novel in a number of ways.
We create probabilistic information about the full exoplanet population
\emph{not} by weighting the detected objects by the inverse detection
efficiency (\citealt{dong, petigura}), but rather by forward modeling the
observed distribution subject to the detection efficiency.
\hoggtodo{Check that Dong \& Zhu do indeed weight by inverse probability.}
Re-weighting the data by inverse selection probability---a method called
V-max in the quasar and galaxy luminosity function contexts---is not wrong,
but it produces a higher-variance estimate of the population than a justified
likelihood approach.
One simple intuition along these lines is that the inverse probability weights
can be large in dynamic range, so the effective number of samples, after
weighting, can be much smaller than the actual number of samples.
Another is that it tends to upweight most strongly the most marginal
detections in the data sample, since these tend to be the samples that lie in
the region of lowest completeness.  In this work, we build a justified
likelihood of the observed data, taking into account the completeness function
or detection efficiency.
Our likelihood function is a variable-rate Poisson likelihood; it makes the
fundamental assumption that the data points are independently drawn from the
model, but is agnostic about all other properties of the data or distribution
functions.
Indeed, the assumptions behind this study are weaker than those of any
previous study of the radius and period distribution of the Earth-analog
population (\citealt{dong, petigura}).

@@

We have made the first measurement of the Earth-like exoplanet population that
is based on forward modeling, takes proper account of the observational
uncertainties on planet radii, and makes use of extremely flexible
non-separable distributions for radius and period.
We find \hoggtodo{BLAH BLAH BLAH.}
